---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# drlearner

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/christophergandrud/dflearner/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/christophergandrud/dflearner/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Doubly robust machine learner (DR Learner) for Heterogeneous Treatment Effect 
Estimation from [Kennedy (2022)](https://arxiv.org/pdf/2004.14497.pdf) to enable 
approximately optimal policy targeting using best linear projections.

## Installation

You can install the development version of blpopt from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("christophergandrud/blpopt")
```

## About

Imagine we want to target individuals in a population with a treatment. The 
effect of the treatment is heterogeneous, i.e. it is conditional on the 
individuals' characteristics (CATE). However, we don't fully observe the characteristics
of each individual. How might we make optimal treatment decisions based 
on the characteristics we do observe?

This package is one step in the following proposed process for to make 
approximately optimal targeting decisions:

1. Estimate CATE, e.g. using `dr_learner` from 
this package or `causal_forest` from the [grf package](https://grf-labs.github.io/grf/).

2. Find the best linear projection of the CATE given the characteristics of the 
individuals we do observe. You can do this using the [blpopt package](https://github.com/christophergandrud/blpopt).

### Formal setup

Imagine data $Z = (X,W,Y)$

$$
\begin{aligned}
\pi(x) & =\mathbb{P}(W=1 \mid X=x) \\
\mu_a(x) & =\mathbb{E}(Y \mid X=x, W=a) \\
\eta(x) & =\mathbb{E}(Y \mid X=x)
\end{aligned}
$$

The aim of conditional average treatment effect estimation is to estimate the difference of the regression functions under treatment $(W = 1)$ versus control $(W = 0)$

$$
\tau(x) \equiv \mu_1(x)-\mu_0(x).
$$

### Doubly Robust Learner (DR Learner)

[Kennedy (2022, 10)](https://arxiv.org/pdf/2004.14497.pdf) proposes using sample slitting estimate the conditional average treatment effects using the following algorithm:

Let $(D^n_1,D^n_2$) be two independent samples of $n$ observations of the data mentioned in the set up.

*Step 1. Nuisance training*

- Construct estimates of $\widehat{\pi}$ of the propensity scores $\pi$ using $D^n_1$.
- Construct estimates $(\widehat{\mu}_{0},\widehat{\mu}_{1})$ of the regression functions $(\mu_0,\mu_1)$ using $D^n_1$.

*Step 2. Pseudo-outcome regression:*

$$
\widehat{\varphi}(Z)=\frac{W-\widehat{\pi}(X)}{\widehat{\pi}(X)\{1-\widehat{\pi}(X)\}}\\{\{Y-\widehat{\mu}_A(X)\\}\}+\widehat{\mu}_1(X)-\widehat{\mu}_0(X)
$$

and regress it on covariates $X$ in the test sample $D^n_2$, giving us

$$
\widehat{\tau}_{d r}(x)=\widehat{\mathbb{E}}_n\\{\widehat{\varphi}(Z) \mid X=x\\}
$$

*Step 3. Estimate outcome given covariates*

Not included in Kennedy (2022), but needed for the best linear projection: 
estimate the outcome given the covariates $X$ using sample $D^n_2$.

*Step 4. Best linear projection of the CATE*

Using the output of the DR Learner procedure above, now find the best
linear projection of the CATE given observed covariates $A$ where $A \subseteq X$.

## Example

```{r}
xfun::pkg_attach2("blpopt", "drlearner", "tidyr", "ggplot2")
theme_set(theme_minimal())

# Simulate data
set.seed(3214)
n <- 2000
p <- 20
X <- matrix(rnorm(n * p), n, p)

# CATE varies along one dim only.
tau_ex <- function(x) {
  1 / (1 + exp(-x))
}
TAU <- tau_ex(X[, 3])

# Propensity  and Outcome vary along 2 and 5 dimensions only.
W <- rbinom(n, 1, 1 / (1 + exp(-X[, 1] - X[, 2])))
Y <- pmax(X[, 2] + X[, 3], 0) + rowMeans(X[, 4:6]) / 2 + W * TAU + rnorm(n)

# CATE-BLP with DR Learner (Kennedy 2022)
drl <- dr_learner(X, Y, W)
blp_drl <- cate_blp(drl, X[, 3])
blp_drl_pred <- tibble(blp_drl = predict(blp_drl)$predicted, A = X[, 3])

ggplot(blp_drl_pred, aes(A, blp_drl)) +
    geom_function(fun = tau_ex, col = "black", linetype = "dashed") +
    geom_point(color = "green") +
    ylab("Predicted Treatment Effect\n") + xlab("\nObserved Covariate Values") +
    ggtitle("[1 Simulation] Compare predicted treatment effects from\nCATE-BLP and plugin linear regression ",
    subtitle = "Black dashed line represents the true effect.\nGreen dots are CATE-BLP from DR Learner")
```
